---
title: "Exploratory Coyote Analysis"
author: "K.Rivera \nkarivera@uri.edu"
date: "`r format(Sys.Date(), '%B %dth, %Y')`"
output:
  html_document:
    toc: yes
    theme: yeti
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, # run the code chunks FALSE/TRUE
                      echo=TRUE, # print the source code FALSE/TRUE
                      error=TRUE, # print error messages FALSE/TRUE
                      message=FALSE, # print any message FALSE/TRUE
                      warning=FALSE, # print warning message FALSE/TRUE
                      cache=FALSE) # only run the updated code chunks FALSE/TRUE 

# other options: results="hide" (don't show the results), fig.show="all" or c(1,3) (show all figures or show figures 1 and 3), fig.width (figure width), fig.height (figure height)
# include = FALSE run the code but the source code and the results are not shown
```

# Coyotes in Rhode Island: An Introduction
## As one of the most adaptable apex carnivores, coyotes have stirred contention across human dominated landscapes. In Rhode Island, coyote-human relationships have been vastly unexplored despite local’s high interest in their presence and activity. Information on human-coyote interactions as well as what drives people’s value of coyotes may be an important tool for managing coyotes and people’s relationship with coyotes in the state. Through an online survey, I gathered data on people’s demographics, relationship and interactions with coyotes, as well as other variables which may be important in predicting people’s value of coyotes.
# Hypothesis: 
## 1) Older individuals will have more negative view of coyotes
## 2) Individuals with more education will have a more positive view of coyotes
## 3) Pet owners with pet-coyote interactions will have a more negative view of coyotes
## 4) Livestock owners with livestock-coyote interactions will have a more negative view of coyotes


# Setting up for data cleaning
```{r}
# clean Up R--------------------------------------------------------------------
rm(list=ls())
gc() #garbage collection

# Libraries---------------------------------------------------------------------
library(tidyverse) # data manipulation (includes ggplot and forcats)
library(lubridate) # dates 
library(dplyr)
library(psych)
library(janitor)
library(corrr)
```

# Read in raw data and rename columns; this is mostly categorical data that was collected using 'Qualtrics Survey Software'
```{r}
# Read in data------------------------------------------------------------------
setwd("C:/Users/Kim Rivera/Documents/Coyotes/Questionnaire Data/live survey/r_data")
coyote_raw = read_csv("Coyote Questionnaire_final_dataset.csv") # OG survey data

# Organize data-----------------------------------------------------------------
coyote <- coyote_raw %>% mutate(date = round_date(mdy_hm(`End Date`))) %>% 
  select(-c(1,2,4,7,8,10:13, 16, 17, 158:162)) 
coyote<- coyote[-c(1:12),] # 'Official' survey started Oct 6

## Rename columns---------------------------------------------------------------
names <- colnames(coyote)
colnames(coyote) <- c("response.type", "progress","time","ID","lat","lon","resident","years.residency","not.resident",
                      "not.resident.others","months.RI", "months.RI.text","county","town","town.text",
                      "street","street.text","outdoor","outdoor.text","nature.person","image.relation.nature",
                      "important.ecosystem","beneficial.humans", "beneficial.wildlife","important.hunters","risk.humans","risk.pets",
                      "negative.wildlife","diseases","reaction","reaction.text","fearful","problem.when",
                      "problem.when.text","coyote.population","all.removed","problem.removed","problem.relocated",
                      "all.sterilized","problem.sterilized","preventive.measures","more.research",
                      "education.program","footholds","native","ri.past.10.years","litter.food",
                      "nocturnal","carnivores","10.human.rabies.year","furbearers","NAMWC","place.coyotes",
                      "place.should","fq.hear","fq.see","fq.see.pups","fq.see.roadkill","fq.see.scat",
                      "fq.see.neighbour","fq.see.eat", "fq.aggressive","fq.fed","pet.owner","pet","pet.text","risk.outdoor.cat",
                      "encounter.outdoor.cat","attack.outdoor.cat","missing.outdoor.cat","kill.outdoor.cat",
                      "risk.large.dog","encounter.large.dog","attack.large.dog","missing.large.dog",
                      "kill.large.dog","risk.small.dog","encounter.small.dog","attack.small.dog",
                      "missing.small.dog","kill.small.dog","risk.pet","encounter.pet","attack.pet",
                      "missing.pet","kill.pet","livestock.owner","livestock","livestock.text",
                      "risk.goat","encounter.goat","attack.goat","missing.goat","kill.goat",
                      "risk.sheep","encounter.sheep","attack.sheep","missing.sheep","kill.sheep",
                      "risk.poultry","encounter.poultry","attack.poultry","missing.poultry","kill.poultry",
                      "risk.pig","encounter.pig","attack.pig","missing.pig","kill.pig","risk.cattle",
                      "encounter.cattle","attack.cattle","missing.cattle","kill.cattle","risk.other",
                      "encounter.other","attack.other","missing.other","kill.other",
                      "protect.from.coyote","measures","measures.text","take.care","not.to.worry",
                      "human.ingenuity","plenty.resources","equal.rights","laws.of.nature","humans.rule","delicate.balance",
                      "envt.cata","age","ethnicity","ethnicity.text","education","household.income",
                      "gender","share","seek.info","contact.envt.mgt.nat.resources","contact.police","contact.uri",
                      "contact.wildlife.center","contact.nature.center","contact.animal.control",
                      "contact.envt.mgt.law.enforce","date")
```

## Since this data is mostly categorical, let's convert character columns to factors and create a new column, 'flagged' for data points (participants) that do not meet survey standards
```{r}
# Convert all characters to factors, can change back to numeric, etc later if needed
coyote = coyote %>% mutate_if(sapply(coyote, is.character), as.factor)

# create new column which ID's unusable data, 0 = good data, 1's = flagged 
# start with survey.type and flag 'Survey Preview' and 'Spam'
coyote = coyote %>% mutate(flagged = ifelse(response.type == "Spam", 1,
                                            ifelse(response.type == "Survey Preview", 1, 0)))
# Now remove these from our cleaned dataset
coyote = filter(coyote, flagged != 1)
```

## Let's go through our variables of interest (and realted variables) and clean up the data
### Age
```{r}
## Check Out Age Distribution
ggplot(coyote, aes(x=age)) + geom_histogram(fill ='yellow4', color = 'yellow4') +
  labs(x = "Age", y = "Counts", title = "Participant Age Distribution") +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=10))
  
# Hmm, I believe it is unlikely a participant is over 110 y/o. Either this was entered wrong or perhaps the 
# participant doesn't want to share their age, or did not take this survey seriously. Let's look at the data 
# point and investigate other responses. 
view = coyote %>% filter(age > 100) # it looks like other data is legitimate, we will therefor change age to NA
coyote <- coyote %>% mutate(age = na_if(age, 120))
```

### Gender
```{r}
levels(coyote$gender)

# Let's convert unusable entries to NA
coyote <- coyote %>% mutate(gender = recode_factor(gender, "Apache attack helicopter" = NA_character_)) %>% 
  mutate(gender = recode_factor(gender, "mmm, no." = NA_character_)) %>% 
  mutate(gender = recode_factor(gender, "I don't believe in gender" = NA_character_)) %>%
  mutate(gender = recode_factor(gender, "Bi curious Buffalo." = NA_character_))

# collapse and correct level labels
coyote <- coyote %>% mutate(gender = fct_collapse(gender, 
                                                  male = c("m", "M", "Male", "boy", "MALE","Make", "Dude",
                                                           "male.  However it is interesting that the previous question offered selections but this question did not.",
                                                           "Man", "masculine", "male"),
                                                  female = c("Femail", "More or less female", "femal", "woman", "Woman", "F", "f", "Female", "female", "is that still a thing? female"),
                                                  non.binary = c("Female, non-binary/fluid" ,"non binary", "Genderqueer/Nonbinary", "Non-Binary","Non-binary","non-binary")))

## Check Out Distribution
ggplot(coyote, aes(x= age, fill = gender)) + geom_bar() +
  labs(title = 'Participant Gender Distribution by Age')+
  scale_x_continuous(breaks = seq(10, 100, by = 5))+
theme_minimal()
```

### Education
```{r}
levels(coyote$education)
# relevel education
coyote <- coyote %>% mutate(education = factor(education)) %>%
  mutate(education = fct_relevel(education, c("Elementary School", "Middle School","High School",  "Technical or vocational degree", "2-year college degree or certificate", "4-year college Degree", "Graduate Degree")))

ggplot(coyote, aes(education, fill = education)) +
  geom_bar() +
  labs(x = "Education", y = "Counts", title = "Level of Education in RI") + 
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 
```

### Household Income
```{r}
# relevel income
coyote <- coyote %>% mutate(household.income = factor(household.income)) %>%
  mutate(household.income = fct_relevel(household.income, c("Under $15,000", "Between $15,000 and $29,999","Between $30,000 and $49,999", "Between $50,000 and $74,999","Between $75,000 and $99,999","Between $100,000 and $149,999","Between $150,000 and $199,999", "$200,000 or more" )))

ggplot(coyote, aes(household.income, fill = household.income)) +
  geom_bar() +
  labs(x = "Household Income", y = "Counts", title = "Household Income RI") + 
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 

# Let's see if income is related to education
ggplot(coyote, aes(household.income, fill = education)) +
  geom_bar() +
  labs(x = "Household Income", y = "Counts", title = "Education by Household Income") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 
```

### County
```{r}
levels(factor(coyote$county))

coyote <- mutate(coyote,county = replace(county, county=="I do not wish to share this information","Not_shared"))

# get the levels ordered by counts 
levels <- coyote %>% 
  group_by(county) %>% 
  summarize(count=n()) %>% 
  arrange(count) %>% 
  select(county) %>% 
  as_vector()

coyote %>% 
  mutate(county = factor(county, levels = levels)) %>% 
  ggplot(aes(x = county, fill = county)) + 
  geom_bar() + 
  labs(x = "County", y = "Counts", title = "Counts per county") + 
  theme_minimal() + 
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 

# Let's see if county is related to education
ggplot(coyote, aes(county, fill = education)) +
  geom_bar() +
  labs(x = "County", y = "Counts", title = "Education by County") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 
```

### Pet Demographics
```{r}
coyote <- coyote %>% 
  mutate(pet.owner = ifelse(str_detect(pet.owner,"(?i)yes"),1,0))
         
# correct non-informative levels to NAs
coyote <- coyote %>% mutate(pet = recode_factor(pet, "Tamed Coyote" = NA_character_)) %>% 
  mutate(gender = recode_factor(pet, "Two" = NA_character_)) %>% 
  mutate(gender = recode_factor(pet, "I have 9 indoor pets of various sizes and species" = NA_character_))%>% 
  mutate(gender = recode_factor(pet, "All indoors and never unattended" = NA_character_))

# Make selection wider data
coyote <- coyote %>% 
  mutate(pet.outdoor.cat = ifelse(str_detect(pet,"(?i)Outdoor.*Cat"),1,0),
         pet.small.dog = ifelse(str_detect(pet,"(?i)Small.*dog"),1,0),
         pet.lrg.dog = ifelse(str_detect(pet,"Large.*dog"),1,0))

# Change data from factor to character for mutate and case_when
coyote <- coyote %>% 
  mutate(pet = as.character(pet)) 
class(coyote$pet)

coyote <- coyote %>% 
  mutate(pet.text = as.character(pet.text)) 
class(coyote$pet.text)

# categorize the input text from participants
coyote <- coyote %>% 
  mutate(pet.text = case_when(pet.text %in% c("2 cats","cats (indoor only)",
                                              "i am active in cat rescue and have previously had outdoor feral cats. I moved them indoors.", 
                                              "Lost a cat to coyotes now indoor only") ~ "indoor cat",
                              pet.text %in% c("2 dogs","my dog is 45#, which I consider medium sized.") ~ "large dogs",
                              pet.text %in% c("3cats, 1. Corgi") ~ "outdoor cat, Corgi",
                              pet.text %in% c("cat goes out only if I am in yard") | str_detect(pet.text,"(?i)indoor/outdoor|cat.*night") ~ "int/out cat",
                              TRUE ~ pet.text)) %>% 
  mutate(pet.indoor.cat = ifelse(str_detect(pet.text,"(?i)indoor.*cat|house.*cat|door.*cat|indoor.*animal|inside.*cat"),1,0),
         pet.outdoor.cat = replace(pet.outdoor.cat,str_detect(pet.text,"(?i)outdoor.*cat|stray.*cat|out.*cat"),1),
         pet.small.dog = replace(pet.small.dog,str_detect(pet.text,"(?i)Corgi|small.*dog"),1),
         pet.lrg.dog = replace(pet.lrg.dog,str_detect(pet.text,"(?i)lab|Medium-sized.*dog|lbdog|Border|large.*dog"),1),
         pet.bird = ifelse(str_detect(pet.text, "(?i)bird|macaw|parrot|cockatoo"),1,0),
         pet.rabbit = ifelse(str_detect(pet.text, "(?i)\\brabbit\\b|house.*rabbit|bunny|indoor.*rabbits"),1,0),
         pet.reptile = ifelse(str_detect(pet.text, "(?i)reptile|snake|herp|lizard|tortoise|turtle"),1,0),
         pet.ferret = ifelse(str_detect(pet.text, "(?i)ferret"),1,0), 
         pet.rodent = ifelse(str_detect(pet.text, "(?i)gerbil|guinea.*pig|hamter"),1,0))

# Lets check out population of pet owners
levels <- coyote %>% 
  pivot_longer(pet.outdoor.cat:pet.rodent, names_to="index",values_to="value") %>% 
  filter(value == 1) %>% 
  group_by(index) %>% 
  summarize(count=n()) %>% 
  arrange(count) %>% 
  select(index) %>% 
  as_vector()

# data with including others answers 
coyote %>% 
  pivot_longer(pet.outdoor.cat:pet.rodent, values_to="value",names_to="index") %>%
  filter(value==1) %>% 
  mutate(index = factor(index, levels=levels)) %>% 
  ggplot(aes(x = index, fill = index)) + 
  geom_bar() + 
  labs(x = "Pets", y = "Counts", title = "Pet ownership") + 
  theme_minimal() + 
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15))
```

### Pet-Coyote Interactions 
#### This is defined as people experiencing: an encounter between their pet and a coyote, an attack on their pet by a coyote or a death of pet by a coyote

```{r}
# Create new column for having had any pet-coyote interaction (only focus on outdoor cats and dogs for now)
pet.int = coyote %>%
  select(encounter.outdoor.cat:kill.small.dog ) %>% # select all columns about pet interactions with coyotes
  select(-contains("risk")) %>%
  mutate_all(~case_when(. == "Yes" ~ 1,
                        . == "No" ~ 0)) %>% # change yes/no for 1/0
  add_column(pet.int = rowSums(., na.rm=TRUE)) %>% # sum all the interactions 
  mutate(pet.int = ifelse(pet.int > 0,"Yes","No")) # # if more than one interaction, assign Yes 

## Add the column to the dataset 
coyote$pet.int <- pet.int$pet.int 
coyote <- coyote %>% 
  mutate(pet.int = replace(pet.int, pet.owner == "No", NA))
# Change pet.int from yes/no, to 1/0
coyote = coyote %>%
  mutate(pet.int = case_when(
    pet.int == "Yes" ~ 1,
    pet.int == "No" ~ 0))
```

### Livestock Demographics
```{r}
# Make selection wider data
coyote <- coyote %>% 
  mutate(livestock.cattle = ifelse(str_detect(livestock,"(?i)cattle"),1,0),
         livestock.goat = ifelse(str_detect(livestock,"(?i)goat"),1,0),
         livestock.poultry = ifelse(str_detect(livestock,"(?i)poultry"),1,0),
         livestock.pig = ifelse(str_detect(livestock,"(?i)pig"),1,0),
         livestock.sheep = ifelse(str_detect(livestock,"(?i)sheep"),1,0))

coyote <- coyote %>% 
  mutate(livestock.bee = ifelse(str_detect(livestock.text, "(?i)bees"),1,0),
         livestock.camelid = ifelse(str_detect(livestock.text, "(?i)llama|alpaca"),1,0),
         livestock.poultry = replace(livestock.poultry,str_detect(livestock.text,"(?i)peacock|hen"),1),
         livestock.equine = ifelse(str_detect(livestock.text,"(?i)horse|donkey"),1,0),
         livestock.cattle = replace(livestock.cattle,str_detect(livestock.text,"(?i)cattle|goat"),1),
         livestock.shellfish = ifelse(str_detect(livestock.text, "(?i)quoahog"),1,0),
         livestock.rabbit = ifelse(str_detect(livestock.text, "(?i)outdoor.*rabbits|rabbits"),1,0))

# counts by levels 
levels <- coyote %>% 
  pivot_longer(livestock.cattle:livestock.rabbit, names_to="index",values_to="value") %>% 
  filter(value == 1) %>% 
  group_by(index) %>% 
  summarize(count=n()) %>% 
  arrange(count) %>% 
  select(index) %>% 
  as_vector()

# all data 
coyote %>% 
  pivot_longer(livestock.cattle:livestock.rabbit, names_to="index",values_to="value") %>% 
  mutate(index = factor(index, levels = levels)) %>% 
  filter(value == 1) %>% 
  ggplot(aes(x = index, fill = index)) + 
  geom_bar() + 
  labs(x = "Livestock Ownership", y = "Counts", title = "Livestock Ownership") + 
  theme_minimal() + 
  theme(legend.position="none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15)) 
```

### Livestock-Coyote Interactions 
#### This is defined as people experiencing: an encounter between their livestock and a coyote, an attack on their livestock by a coyote or a death of livestock by a coyote

```{r}
## Create new column for having had any pet-coyote interaction (only focus on outdoor cats and dogs for now)
livestock.int = coyote %>%
  select(encounter.goat:kill.cattle) %>% # select all columns about livestock interactions with coyotes
  select(-contains("risk")) %>%
  mutate_all(~case_when(. == "Yes" ~ 1,
                        . == "No" ~ 0)) %>% # change yes/no for 1/0
  add_column(livestock.int = rowSums(., na.rm=TRUE)) %>% # sum all the interactions 
  mutate(livestock.int = ifelse(livestock.int > 0,"Yes","No")) # # if more than one interaction, assign Yes 

## Add the column to the dataset 
coyote$livestock.int <- livestock.int$livestock.int 
coyote <- coyote %>% 
  mutate(livestock.int = replace(livestock.int, pet.owner == "No", NA))
# Change livestock.int from yes/no, to 1/0
coyote = coyote %>%
  mutate(livestock.int = case_when(
    livestock.int == "Yes" ~ 1,
    livestock.int == "No" ~ 0))
```

### Human-Coyote Relationships
##### Lets focus on participants thoughts on coyotes, e.g. ranking of benefits and risks of coyotes
```{r}
# Starting with benefits strong disagree [1] -> strong agree [5]
coyote <- coyote  %>%
  mutate(important.ecosystem = fct_relevel(important.ecosystem, c("Strongly Disagree", "Disagree", "Neither agree nor disagree",  "Agree", "Strongly Agree")))
coyote <- coyote  %>%
  mutate(beneficial.humans = fct_relevel(beneficial.humans, c("Strongly Disagree", "Disagree", "Neither agree nor disagree",  "Agree", "Strongly Agree")))
coyote <- coyote  %>%
  mutate(beneficial.wildlife = fct_relevel(beneficial.wildlife, c("Strongly Disagree", "Disagree", "Neither agree nor disagree",  "Agree", "Strongly Agree")))
coyote <- coyote  %>%
  mutate(important.hunters = fct_relevel(important.hunters, c("Strongly Disagree", "Disagree", "Neither agree nor disagree",  "Agree", "Strongly Agree")))

# Now risks strong disagree [5] -> strongly agree [1]
coyote <- coyote  %>%
  mutate(risk.humans = fct_relevel(risk.humans, c("Strongly Agree", "Agree", "Neither agree nor disagree", "Disagree", "Strongly disagree")))
coyote <- coyote  %>%
  mutate(risk.pets = fct_relevel(risk.pets, c("Strongly Agree", "Agree", "Neither agree nor disagree",  "Disagree", "Strongly disagree")))
coyote <- coyote  %>%
  mutate(negative.wildlife = fct_relevel(negative.wildlife, c("Strongly Agree", "Agree", "Neither agree nor disagree",  "Disagree", "Strongly disagree")))
coyote <- coyote  %>%
  mutate(diseases = fct_relevel(diseases, c("Strongly Agree", "Agree", "Neither agree nor disagree",  "Disagree", "Strongly disagree")))
```

# Coyote Data Analysis
## We first use a Cronbach’s Alpha to determine the correlation between all questions used to assess participant’s value of coyotes. This is done by unclassifying factorized Likert data. The resulting alpha allows us to determine the reliability of multi-item scales 

```{r}
coyote.attitude = coyote %>%
  select(c(important.ecosystem, beneficial.humans, beneficial.wildlife, important.hunters,
           risk.humans, risk.pets, negative.wildlife, diseases)) %>%
  mutate_at(c("important.ecosystem", "beneficial.humans", "beneficial.wildlife", "important.hunters",
              "risk.humans", "risk.pets", "negative.wildlife", "diseases"), ~unclass(.))

# This test will tell us how well items here correlate
# ideally want an alpha coefficient of 0.08 or higher for short lists of items 
psych::alpha(coyote.attitude)
# this gives us a raw alpha of .84, a .8 correlation is considered sufficient for short lists
```

## Kendall's Rank Coefficient 
### Kendall’s tau coefficient is a non-parametric correlation test used to determine the relationship between individual value questions and our hypothesis. Kendall’s rank coefficient uses ranked pairs to determine the association direction (positive or negative) and strength (stronger as tau approaches ±1). 
```{r}
# Let's look at our hypothesis and start by analyzing the ordinal level data, education
# Hyp 2) Individuals with more education will have a more positive view of coyotes

kendall = coyote %>% select(c(beneficial.humans, beneficial.wildlife, important.ecosystem, important.hunters, 
                              diseases, negative.wildlife, risk.humans, risk.pets, education)) %>%
  mutate_all(~unclass(.))


eco.ed = cor.test(kendall$important.ecosystem, kendall$education, method="kendall")
ben.hum.ed = cor.test(kendall$beneficial.humans, kendall$education, method="kendall")
ben.wild.ed = cor.test(kendall$beneficial.wildlife, kendall$education, method="kendall")
hunt.ed = cor.test(kendall$important.hunters, kendall$education, method="kendall")
dis.ed = cor.test(kendall$diseases, kendall$education, method="kendall")
neg.wild.ed = cor.test(kendall$negative.wildlife, kendall$education, method="kendall")
risk.hum.ed = cor.test(kendall$risk.humans, kendall$education, method="kendall")
risk.pet.ed = cor.test(kendall$risk.pets, kendall$education, method="kendall")

eco.ed
ben.hum.ed 
ben.wild.ed
hunt.ed 
dis.ed 
neg.wild.ed 
risk.hum.ed 
risk.pet.ed

# Hyp 3) Pet owners pet-coyote interactions will have a more negative view of coyotes

eco.pet = cor.test(kendall$important.ecosystem, coyote$pet.int, method="kendall")
ben.hum.pet = cor.test(kendall$beneficial.humans, coyote$pet.int, method="kendall")
ben.wild.pet = cor.test(kendall$beneficial.wildlife, coyote$pet.int, method="kendall")
hunt.pet = cor.test(kendall$important.hunters, coyote$pet.int, method="kendall")
dis.pet = cor.test(kendall$diseases, coyote$pet.int, method="kendall")
neg.wild.pet = cor.test(kendall$negative.wildlife, coyote$pet.int, method="kendall")
risk.hum.pet = cor.test(kendall$risk.humans, coyote$pet.int, method="kendall")
risk.pet.pet = cor.test(kendall$risk.pets, coyote$pet.int, method="kendall")

eco.pet
ben.hum.pet
ben.wild.pet
hunt.pet
dis.pet
neg.wild.pet
risk.hum.pet
risk.pet.pet

# Hyp 4) Pet owners livestock-coyote interactions will have a more negative view of coyotes

eco.livestock = cor.test(kendall$important.ecosystem, coyote$livestock.int, method="kendall")
ben.hum.livestock = cor.test(kendall$beneficial.humans, coyote$livestock.int, method="kendall")
ben.wild.livestock = cor.test(kendall$beneficial.wildlife, coyote$livestock.int, method="kendall")
hunt.livestock = cor.test(kendall$important.hunters, coyote$livestock.int, method="kendall")
dis.livestock = cor.test(kendall$diseases, coyote$livestock.int, method="kendall")
neg.wild.livestock = cor.test(kendall$negative.wildlife, coyote$livestock.int, method="kendall")
risk.hum.livestock = cor.test(kendall$risk.humans, coyote$livestock.int, method="kendall")
risk.pet.livestock = cor.test(kendall$risk.pets, coyote$livestock.int, method="kendall")

eco.livestock
ben.hum.livestock
ben.wild.livestock
hunt.livestock
dis.livestock
neg.wild.livestock
risk.hum.livestock
risk.pet.livestock
```
#### It's clear, that looking at these 'attitude' scale questions individually, there are very weak positive and negative correlations. The strongest correlations are: risk to humans with pet interaction, risk to pets with pet interactions, and risk to humans with livestock interactions. All of these are negativelty correlated meaning as people have more pet and livestock interactions with coyotes, they view coyotes more negatively (because the 'attitude' scale goes form negative --> positive) 


## Ordinal Regression
### I use the median from each participant’s responses as a continuous responsive variable for the benefits and another for the risks to conduct an ordinal logistic regression analysis 

```{r}
# first organize the data and take the mean
kendall = kendall %>% rowwise() %>%
  mutate(coyote.attitude = mean(c(important.ecosystem, beneficial.humans, beneficial.wildlife, important.hunters)), na.rm = TRUE) %>%
  select(-na.rm)

# add this data back to original dataset
coyote$coyote.attitude <- kendall$coyote.attitude 

# tell R how these attitudes rank
coyote <- coyote %>%
  mutate(coyote.attitude = ordered(coyote.attitude, levels = c("1", "1.5", "2","2.5","3","3.5","4","4.5","5"))) 

# This library is for ordinal regression analysis
library(MASS)
```

### Let's run a few models 
```{r}
# This includes all of our variables of interest
mod1 = polr(coyote.attitude ~ age + education + pet.int + livestock.int, data = coyote, method="logistic", Hess=TRUE)
AIC(mod1)

#Let's try variations of these variables and compare AIC's
mod2 = polr(coyote.attitude ~ education + pet.int + livestock.int, data = coyote, method="logistic", Hess=TRUE)
AIC(mod2)

mod3 = polr(coyote.attitude ~ age + pet.int + livestock.int, data = coyote, method="logistic", Hess=TRUE)
AIC(mod3)

mod4 = polr(coyote.attitude ~ pet.int + livestock.int, data = coyote, method="logistic", Hess=TRUE)
AIC(mod4)

AIC(mod1, mod2, mod3, mod4)
```

#### It appears the first model is our best fit, mod1, let's explore that more
```{r}
summary(mod1)
coef(mod1) #extract model coef 
ctable = coef(summary(mod1))
p = pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2  ## add the p-values to the table
ctable = cbind(ctable, "p value" = p)
ctable
```

### Interpretting our model
```{r}
ci <- confint(mod1)
exp(cbind(odd_ratio = coef(mod1), ci))
```
#### This code provides the proportional odds ratio. This ratio allows us to interpret the odds participants will have a more positive or negative attitude of coyotes. Here we see that with increasing age and interactions between coyotes and livestock, participants odds of having a positive attitude towards coyotes decreases. Education varibales are harder to interpret although all odds are above 1, meaning they are more likely to have a postive attitude towards coyotes given education than the other variables (age and interactiosn betwen coyotes and pets & livestock)

# Exploratory Data Analysis
## We can also plot our variables of interest and see if we can find any obvious relationships between our independent variables of interest and peoples attitutes towards coyotes
```{r}
coyote %>% filter(is.na(age)== FALSE) %>%
  ggplot(aes(x = age, y = coyote.attitude, fill = coyote.attitude))+
  geom_jitter(color = "darkgray", size = .3, alpha =.9)+
  geom_boxplot()

ggplot(coyote, aes(education, fill = coyote.attitude)) +
  geom_bar() +
  labs(x = "Education", y = "Counts", title = "Coyote Attitude by Education") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15))

ggplot(coyote, aes(pet.int, fill = coyote.attitude)) +
  geom_bar() +
  labs(x = "Pet Interaction", y = "Counts", title = "Coyote Attitude by Pet Interaction") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15))

ggplot(coyote, aes(livestock.int, fill = coyote.attitude)) +
  geom_bar() +
  labs(x = "Livestcok Interaction", y = "Counts", title = "Coyote Attitude by Livestock Interaction") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(family = "LM Roman 10 Bold", size=15))
```
